---
title: "MuPETFlow-introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MuPETFlow-introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

**MuPET-Flow** is a graphical user interface (GUI) tool designed to estimate ploidy from DNA cells stained with fluorescent dyes and analyzed via flow cytometry. It automates the processes of file uploading and configuration, peak fluorescence intensity detection, multiple histogram visualizations, peak error curation, ploidy and genome size calculations, and result exportation.

```{r setup}
library(MuPETFlow)
```

## How to run MuPET-Flow

Once **MuPET-Flow** is installed, you can run the application using the following function:

```r
runMuPETFlow()
```
Note: The first time you launch the tool, it will attempt to install additional **Bioconductor** dependencies that are not managed via `install.packages()`. This process might take a few minutes.

## Description

Once the application is launched, you can either:

1. Load your experimental data.
2. Run an in-app example by clicking the 'Example' button.

For the first case, selecting the channel where the data was acquired is mandatory. For the second case, the example channel `FL4-A` is automatically detected for demonstration purposes.

After launching the app, you can follow the app flow, which is divided into three tabs: _**Peaks**_, _**Regression**_ and _**Summary**_.  Below is a general description of the options available in each tab:

### Peaks
* **Select a sample (optional):** Allows visual exploration of individual samples if desired.
* **Adjust smoothing (optional):** Adjusts the histogram curve for noisy samples.
* **Adjust window width (optional):** Defines the interval where the app will look for peaks.
* **Select minimum cell count to call a peak (optional):** Useful for samples with a low number of events.
* **Select maximum number of peaks to plot (optional):** Useful for samples with heterogeneous populations where more peaks are present.

### Regression
* **Select type of analysis:** Choose between "Ploidy" or "Genome size" analysis.
* **Select number of standards:** A minimum of two different standards is required, but more are recommended.
* **Select standard samples and values:** This is the ploidy or genome size of your standards.

### Summary
* **Results preview:** Creates a compiled figure with histograms for all samples.
* **Save plot:** Saves the histograms in either PNG or TIFF format with customizable size and quality. Optionally, you can control the grid layout.
* **Save table:** Exports the parameters used and the estimated ploidy or genome size as a CSV file.

If any errors are detected in the analyzed samples, you can go back to the Peaks tab to review the parameters. Note that the regression must also be re-done after parameter adjustments.
